---
title: "New_ITA_paper"
author: "Carolina Castro-Sanguino"
date: "13/11/2020"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(root.dir= '/tmp',echo = TRUE)

##Need renv package 
#renv::init() ##Creates private library

rm(list=ls())

library(rgdal)
library(rgeos)
library(raster)
library(magrittr)
library(dplyr)
library(maps)
library(lubridate)
library(ggcorrplot)
library(FSSgam)
library(mgcv)
library(ggsn)
library(ape)
library(ggpubr)
library(stringr)
library(viridis)
library(tidyr)
library(gplots)
library(egg)

##The final dataset for modelling: AIMS and Catlin merged with cyclone descriptors and other metrics
load("ita_cover.Rdata")

##Ita track from BOM for plotting
track<-read.csv("Ita_track_BOM.csv")

##All cyclone Descriptors (with those excluded for Appendix)
wavedata<-read.csv("wavedata.csv")

```



# Explore wave data

```{r}


##----------------------------------
##Figure 2: correlated wave metrics
##----------------------------------
wavedata%<>%filter(stid %in% ita_ndf$stid)##Only for selected reefs

p1=ggscatter(wavedata, x = "whmax", y = "wp90",
             conf.int = TRUE,
             cor.coef = TRUE, cor.method = "spearman",color="cornflowerblue",
             xlab = "Max.wave Height (m)", ylab = "Energy flux (Wf)\n90th percentile")+ geom_smooth(color="black")#add = "reg.line",
#
#
p2=ggscatter(wavedata, x = "whmax", y = "wpc",
             conf.int = TRUE,
             cor.coef = TRUE, cor.method = "spearman",color="cornflowerblue", font.label = c(12),
             xlab = "Max.wave Height (m)", ylab = "Cumulative Energy\nflux (kW/m)")+ geom_smooth(color="black")##add = "reg.line",
#
#
p3=ggscatter(wavedata, x = "wh4mH", y = "wpc",
             conf.int = TRUE,
             cor.coef = TRUE, cor.method = "spearman", color="cornflowerblue",
             xlab = "Wave height\nthreshold (4MW)", ylab = "Cumulative Energy\nflux (kW/m)")+ geom_smooth(color="black")##add = "reg.line",

p4=ggscatter(wavedata, x = "wh4mH", y = "Ubms",
             conf.int = TRUE,
             cor.coef = TRUE, cor.method = "spearman", color="cornflowerblue",
             xlab = "Wave height\nthreshold (4MW)", ylab = "Max. near-bed\norbital velocity amplitud")+ geom_smooth(color="black")##add = "reg.line",


ggarrange (p1,p2,p3,p4, ncol=2)
#ggsave(file="Outputs/Wavecorr.png",width=7, height=6,dpi=300)

##-----------------------------------------------------------------------------------
##Transform all potential variables, check correlations and then subset for modelling
##-----------------------------------------------------------------------------------
ita_ndf%<>%left_join(unique(df.1[c(4,5,7)]))#Add coordinates per transect

##Check predictors distributions and transform/center (always transform first, and then center or standardize if needed)
contpreds=names(ita_ndf[-c(1:3)])

# for(p in 1:length(contpreds)){
# par(mfrow=c(1,2))
#  hist(ita_ndf[,(contpreds[p])],main=contpreds[p])
#  plot(jitter(ita_ndf[,contpreds[p]]))
#  }

##LOG: wpc,whmax,4mwh,wpmax,ubms,Th1
##sqrt: Acropora
##sqrt: Obranching

##Transform to normalise

ita_ndf %<>%
  mutate(log.wpc=log(wp.c+1),
         log.4mwh=log(wh.4mH+1),
         log.ubms=log(Ubms+1),
         log.whmax=log(wh.max+1),
         log.T1h=log(T1h+1), 
         log.wpmax=log(wp.max+1),
         sqrt.Acropora=sqrt(Acropora),
         sqrt.Otherbranching=sqrt(Otherbranching),
         sqrt.change=sqrt(Rchange+1),
         log.postIta=log(postIta),
         sqrt.mwind=sqrt(Max_wind),
         ndist=log(distance))



newdat=ita_ndf[,c("Rchange","sqrt.change","reef_name","stid","nc_ubedmean","sqrt.Acropora",
                  "log.T1h","log.ubms","log.wpmax", "log.wpc","log.4mwh","log.whmax","log.postIta", "sqrt.mwind", "ndist")]

#Edit names for figure
newdat%<>%dplyr::rename("nc-Ub"="nc_ubedmean",`% Acropora`= "sqrt.Acropora","Ub-T1h"="log.T1h","Ita-Ub"="log.ubms",`Wf max`="log.wpmax","Wfc"="log.wpc",`4mWh`="log.4mwh","Whmax"="log.whmax",`time post-Ita`="log.postIta",`Wind max`="sqrt.mwind","mindist"="ndist")

cpreds=names(newdat[-c(1:4)])

##Correlate transformed variables (reduced correlation coeff but still need to test independently)
corpred<-round(cor(newdat[,cpreds],use="complete.obs"),2)

##-------------------
##Appendix Figure 1
##-------------------
ggcorrplot(corpred, hc.order = TRUE, type = "lower",lab = TRUE)
#ggsave("Outputs/Correlation_Transformed_FinalPreds.jpeg", device="jpeg",width = 9, height = 6, units = c("in"),dpi = 300)


##Functions to Backtransform response/predictors
square <- function(x){
  return((x**2)-1)
}

loginv <- function(x){
  return(exp(x)-1)
}


```

#Explore best preditors with FSSGAMS
1. Include all predictors and only a metric of %Acroporids
In the presence of community composition measures of orbital velocity are best explaining variability. 

```{r}

##Exclude highly correlated predictors (i.e >0.90 cor)
##wpcum is the least correlated of all--choose and compare RelImp against wh metrics
##remove T1h as highly correlated with log.Ub (which show greater correlation to response)

##NOTE:### Models with excluded variables show they are not important
##test of cumulative wp (RI=0.07)
##test of max wh (RI=0.009)
##test of 4mwh (RI=0.009)

contpreds1=names(ita_ndf[,c("nc_ubedmean","sqrt.Acropora","log.ubms","log.wpc","log.postIta","ndist")])## for modelling
#contpreds1=names(ita_ndf[,c("nc_ubedmean","sqrt.Acropora","log.ubms","log.4mwh","log.postIta","ndist")])## check 4mwh: Not in best model


##Make sure there are at least 2 replicates per reef
countdat=ita_ndf%>%dplyr::select(reef_name,stid)%>%group_by(reef_name)%>%count()
summary(countdat$n)#OK


ita_ndf%<>%
  mutate_if(is.character,as.factor)

usedat=ita_ndf[,c("sqrt.change","reef_name","nc_ubedmean","sqrt.Acropora","log.ubms","log.wpc","log.postIta","ndist")]
#usedat=ita_ndf[,c("sqrt.change","reef_name","nc_ubedmean","sqrt.Acropora","log.ubms","log.4mwh","log.postIta","ndist")]

##is including Acropora important? (model without Acropora)
#usedat=usedat[,c("sqrt.change","reef_name","nc_ubedmean","log.ubms","log.wpc","log.postIta","ndist")]
#contpreds1=names(usedat2[,c("nc_ubedmean","log.ubms","log.wpc","log.postIta","ndist")])#


Mod1=gam(sqrt.change~s(ndist,k=3)+s(reef_name,bs='re'),data=usedat)##Update per model

model.set=generate.model.set(use.dat=usedat,
                             max.predictors=3, 
                             test.fit=Mod1, 
                             k=3, 
                             pred.vars.cont=contpreds1,##Update for each model
                             smooth.smooth.interactions=T,
                             cov.cutoff=0.28, 
                             null.terms="s(reef_name,bs='re')")

mod.list=fit.model.set(model.set)## FITS mod
##extract Variable Importance Scores
var.imp.aic=mod.list$variable.importance$aic$variable.weights.raw

# Model selection table and best models
mod.table=mod.list$mod.data.out
mod.table=mod.table[order(mod.table$AICc),]
less.2AICc=mod.table[which(mod.table$delta.AICc<2),]


View(less.2AICc)

relimp.aic=as.data.frame(var.imp.aic)
relimp.aic$predictor <- rownames(relimp.aic)
rownames(relimp.aic)=relimp.aic$predictor

#edit for graph
relimp.aic$predictor=c("nc-Ub", "%Acropora","Ita-Ub","Ita-Wf","Time post-Ita","min.dist")

##Fit best MOdel
less.2AICc$formula[1]##

mod=gam(sqrt.change~te(sqrt.Acropora,log.ubms, k = 3, bs = c("cr", "cr"))+s(reef_name, bs = "re"), data=usedat,method="REML")##Gaussian OK
summary(mod)#47%


modfit=gratia::appraise(mod)##USE for supplementary
dev.new()
#ggsave("Outputs/BestMod_fit_Tot_rnd.jpeg", modfit,device="jpeg",width = 9, height = 6, units = c("in"),dpi = 300) 


##-----------------------------------------------------------------
##Plot observed total cover change as a function of best predictors
##-----------------------------------------------------------------
##Categorize %Acropora for easier visualization of interaction
classorder=c("<15%","15-35%",">35%")


usedat%<>%mutate(Acropora=NA)%>%
  mutate(Acropora=ifelse(sqrt.Acropora**2 <15, "<15%",
                         ifelse(sqrt.Acropora**2  >=15 & sqrt.Acropora**2 <35.1, "15-35%",">35%")))%>%
  mutate(Acropora=factor(Acropora, levels=classorder))

#----------
##Figure 4
#----------

f1=ggplot(relimp.aic, aes(y=reorder(predictor, -var.imp.aic,sum),var.imp.aic))+geom_col(aes(color=reorder(predictor, -var.imp.aic,sum), fill=reorder(predictor, -var.imp.aic,sum)), alpha=0.4)+scale_colour_hue(h = c(180, 270))+scale_fill_hue(h = c(180, 270))+theme_classic()+theme(legend.position = "none")+ylab("Predictors")+xlab("Relative Importance-AIC")
# ggsave("Outputs/RI_Total.jpeg",device="jpeg",width = 3, height = 4, units = c("in"),dpi = 300)

##Observed data as a fuction of modelled predictors (mod withouth random)
f2=ggplot(usedat, aes(x=log.ubms, y=square(sqrt.change),color=Acropora), size=Acropora)+geom_point(aes(x=log.ubms, y=square(sqrt.change),color=Acropora,size=Acropora), alpha=0.4)+viridis::scale_color_viridis(discrete = TRUE, option = "D")+viridis::scale_fill_viridis(discrete = TRUE)+ylab("Relative change\nin total coral cover")+theme_classic()+xlab("Log Ita-Ub (m/s)")+geom_smooth(aes(color=Acropora,fill=Acropora), method="gam", formula = y~s(x, k=3, bs="cr"))
#ggsave("Outputs/Edited_observed_totalRchange.jpeg",device="jpeg",width = 5, height = 4, units = c("in"),dpi = 300)  

ggarrange(f1,f2, nrow=1, labels=c('a)','b)'))

```

# Explore best Model and the importance of bottom-wave metrics and community composition for total cover  change

```{r}

plot(residuals.gam(mod)~usedat$reef_name)#OK

##check data spatially

spatdat=ita_ndf%>%dplyr::select(c(17,18,3))%>%droplevels()
sp::coordinates(spatdat)<-~lat+lng##convert to spatial df
crs(spatdat)<- CRS("+init=epsg:3857") ##
spatdat2<-spTransform(spatdat, CRS("+init=epsg:32755"))#UTM for estimating spatial autocorrelation
spatdat2$Res=residuals.gam(mod)
sp::bubble(spatdat2,"Res")##No strong spatial pattern in residuals OK

##Moran'sI (value of slope of fitted line between neignouring response values and each coord)
coords=spatdat2@coords
w=fields:::rdist(coords)
Moran.I(x=newdat$sqrt.change, w=w)## -1/(N-1) ## Expected value negative and nearly 0.
##obs close to expected but pvalue=0(reject null of no spatial autocor). 
##pretty small value -0.09 for residual autocorrelation

#---------
##Figure 7
#---------
##Plot ItaUb vs ncUb
p1=ggscatter(usedat, x="log.ubms", y="nc_ubedmean", conf.int = TRUE, cor.coef = TRUE)+geom_smooth(method="gam", formula=y~s(x, bs="cr", k=3), color="grey")+ylab("nc-Ub")+xlab("Ita-Ub")+theme(axis.text = element_text(size=8))

##Plot %Acropora along ncUb
p2=ggplot(usedat, aes(y=nc_ubedmean,x=Acropora, fill=Acropora),color=Acropora)+ylab("nc-Ub")+xlab("Acropora")+geom_violin(alpha=0.7)+viridis::scale_fill_viridis(discrete =TRUE)+theme_classic()+theme(legend.position = "none")+xlab("% Acropora")

ggarrange(p1,p2, nrow=1, labels=c('a)','b)'))
```


2. Run independent models for the change of individual coral groups as a response.
Inspection of changes per coral groups show that Porites Massive was the only coral group that consistently show losses across all reefs (most coral groups show great variations with mean/median of change suggesting no/+ change in cover)


```{r}

##Estimate relative change per Group (Convert each Fg cov to relative cover and calculate change)
pergroupchange=df.1%>%
  mutate(Acropora=((ACR_BRA+ACR_TCD)/Total)*100,
         MSE=((MSE+ACR_OTH)/Total)*100, 
         BRA_nACR=((BRA_nACR+POCI)/Total)*100,
         POR_MASS=((POR_MASS/Total)*100))%>%
  dplyr::select(-c(ACR_BRA,ACR_TCD,POR_TAB,ACR_OTH,POCI,Total))%>%
  tidyr::gather_(key_col="group", value_col="cover", gather_cols=c("Acropora","BRA_nACR","MSE","POR_MASS"))


##Remove sites where Rcover pre-Ita was <5% 

##Acropora
Achange=pergroupchange%>%
  filter(group == "Acropora" & year==2012)%>%
  filter(cover < 5)%>%droplevels()
summary(Achange)

iddel=levels(as.factor(Achange$stid))

Achange=pergroupchange%>%
  filter(group %in% c("Acropora"))%>%
  filter(!stid %in% iddel)%>%
  tidyr::spread(year,cover)%>%
  rename(y2012="2012", y2014="2014")%>%
  mutate(Rchange=(y2014-y2012)/y2012)
##Make sure there are at least 2 replicates per reef
Achange%>%dplyr::select(reef_name,stid)%>%group_by(reef_name)%>%count()
Achange%<>%mutate(transchange=sqrt(Rchange+1))##Transform response

##Other branching(poci and porites)
OBchange=pergroupchange%>%
  filter(group %in% c("BRA_nACR"))%>%
  filter(year==2012 & cover < 5)%>%droplevels()

iddel=levels(as.factor(OBchange$stid))

OBchange=pergroupchange%>%
  filter(group %in% c("BRA_nACR"))%>%
  filter(!stid %in% iddel)%>%
  tidyr::spread(year,cover)%>%
  rename(y2012="2012", y2014="2014")%>%
  mutate(Rchange=(y2014-y2012)/y2012)%>%
  filter(Rchange<4)##Outlier
##Make sure there are at least 2 replicates per reef
OBchange%>%dplyr::select(reef_name,stid)%>%group_by(reef_name)%>%count()
OBchange%<>%mutate(transchange=sqrt(OBchange$Rchange+1))##Transform response

##MSE
MSEchange=pergroupchange%>%
  filter(group %in% c("MSE"))%>%
  filter(year==2012 & cover < 5)%>%droplevels()

iddel=levels(as.factor(MSEchange$stid))

MSEchange=pergroupchange%>%
  filter(group %in% c("MSE"))%>%
  filter(!stid %in% iddel)%>%
  tidyr::spread(year,cover)%>%
  rename(y2012="2012", y2014="2014")%>%
  mutate(Rchange=(y2014-y2012)/y2012)%>%
  filter(Rchange<10)#Outlier
##Make sure there are at least 2 replicates per reef
MSEchange%>%dplyr::select(reef_name,stid)%>%group_by(reef_name)%>%count()
MSEchange%<>%mutate(transchange=log(MSEchange$Rchange+1))##Transform response

##PORMASS (mean change= loss -0.4) This is the only group with mean loss, best for evaluate damage predictors
PORchange=pergroupchange%>%
  filter(group %in% c("POR_MASS"))%>%
  filter(year==2012 & cover < 5)%>%droplevels()

iddel=levels(as.factor(PORchange$stid))

PORchange=pergroupchange%>%
  filter(group %in% c("POR_MASS"))%>%
  filter(!stid %in% iddel)%>%
  tidyr::spread(year,cover)%>%
  rename(y2012="2012", y2014="2014")%>%
  mutate(Rchange=(y2014-y2012)/y2012)%>%
  filter(Rchange<4)#Outlier
##Make sure there are at least 2 replicates per reef
PORchange%>%dplyr::select(reef_name,stid)%>%group_by(reef_name)%>%count()
#Remove Noname and Ribbon 9
PORchange%<>%filter(!reef_name %in% c("No name", "Ribbon 9"))##Check distribution of changes
PORchange%<>%mutate(transchange=sqrt(PORchange$Rchange+1))


##----------------------------------------------
##Combine all and include transformed predictors
##----------------------------------------------

groupchange=rbind(Achange[c(3:8,11:12)],OBchange[c(3:8,11:12)],MSEchange[c(3:8,11:12)],PORchange[c(3:8,11:12)])

groupchange$reef_name<-str_squish(groupchange$reef_name)
groupchange%<>%left_join(ita_ndf[c(3,4,19:24,28,30)], by="stid")

```

Model per group

```{r}


contpreds=names(groupchange[,c("nc_ubedmean","log.ubms","log.wpc","log.postIta","ndist")])
#contpreds=names(groupchange[,c("nc_ubedmean","log.ubms","log.4mwh","log.postIta","ndist")])##Test just in case (4mWh only relevant for Acropora)
#contpreds=names(groupchange[,c("nc_ubedmean","log.ubms","log.whmax","log.postIta","ndist")])##Test just in case (whmax ns)

groupdat=groupchange%>%
  mutate_if(is.character,as.factor)

groupdat=groupdat[,c("group","transchange","reef_name","nc_ubedmean","log.ubms","log.wpc","log.postIta","ndist")]
#groupdat=groupdat[,c("group","transchange","reef_name","nc_ubedmean","log.ubms","log.4mwh","log.postIta","ndist")] 
##4mwh only important for Acropora. Explain as much as Ita-Ub. 
#groupdat=groupdat[,c("group","transchange","reef_name","nc_ubedmean","log.ubms","log.whmax","log.postIta","ndist")]

#Model each coral group
resp.vars=unique(as.character(groupchange$group))

out.all=list()
var.imp=list()
fss.all=list()
top.all=list()

for(i in 1:length(resp.vars)){
  use.dat=groupdat[which(groupdat$group==resp.vars[i]),]
  Mod=gam(transchange~s(ndist,k=3)+s(reef_name,bs='re'),data=use.dat)
  model.set=generate.model.set(use.dat=use.dat,
                               test.fit=Mod, 
                               max.predictors=3, 
                               k=3, 
                               pred.vars.cont=contpreds,
                               smooth.smooth.interactions=T,
                               cov.cutoff=0.28,
                               null.terms="s(reef_name,bs='re')") ##compare
  
  
  out.list=fit.model.set(model.set)
  fss.all=c(fss.all,list(out.list))
  mod.table=out.list$mod.data.out
  mod.table=mod.table[order(mod.table$AICc),]
  out.i=mod.table
  out.all=c(out.all,list(out.i))
  var.imp=c(var.imp,list(out.list$variable.importance$aic$variable.weights.raw))
  all.less.2AICc=mod.table[which(mod.table$delta.AICc<2),]
  top.all=c(top.all,list(all.less.2AICc))
  
}


# Model fits and importance---
resp.vars= c("Acropora","Other Branching","MSE","Porites")
names(out.all)=resp.vars
names(var.imp)=resp.vars
names(top.all)=resp.vars
names(fss.all)=resp.vars

all.mod.fits=do.call("rbind",out.all)
all.var.imp.aic=do.call("rbind",var.imp)
top.mod.fits.aic=do.call("rbind",top.all)


##AIC-based BestMods (Table S3)
top.mod.fits.aic$response=rownames(top.mod.fits.aic)
top.mod.fits.aic %<>%separate(response, c('response', NA), sep="[(.)]")%>% dplyr::select(response, everything())
#write.csv(top.mod.fits.aic[,c("modname","AICc","BIC","delta.AICc","delta.BIC","wi.AICc","wi.BIC","edf","r2.vals")],
#           file="Best_AIC_mods.csv")


#---------
##Figure 5
#---------

heatmap.2(all.var.imp.aic,notecex=0.3,  dendrogram ="none",
          col=colorRampPalette(c("white","lightsteelblue1", "cornflowerblue", "slateblue4"))(10),
          trace="none",key.title = "",keysize=1,
          notecol="black",key=T,
          sepcolor = "black",margins=c(12,15), lhei=c(4,15),Rowv=FALSE,Colv=FALSE)
#ggsave(file="Outputs/Cgroups_RI_AIC_rnd_interactions.jpeg",width = 10, height = 6, units = c("in"),dpi = 300)




##---------------------------------
##Fit best Model (Acropora)
##---------------------------------
top.mod.fits.aic$modname[top.mod.fits.aic$response=="Acropora"]

Ac=groupdat%>%filter(group=="Acropora")%>%droplevels()

mod.ac1=gam(transchange~te(nc_ubedmean, ndist, k = 3, bs = c("cr","cr"))+s(reef_name,bs='re'),dat=Ac,method="REML")
summary(mod.ac1)## 37% rnd model

mod.ac2=gam(transchange~s(log.ubms, k = 3, bs = "cr")+s(reef_name,bs='re'),dat=Ac,method="REML")
summary(mod.ac2)##competing model



##-----------------------------------------------------------------
##Plot observed total cover change as a function of best predictors
##-----------------------------------------------------------------

##Ita Ub
M1=Ac%>%group_by(reef_name)%>%ggplot(aes(x=log.ubms, y=square(transchange)))+
  geom_point(color="forestgreen",alpha=0.5)+geom_smooth(color="forestgreen",method="gam", formula = y~s(x, k=3, bs="cr"))+labs(y=expression("Relative change in"~italic("Acropora")~"cover"))+theme_classic()+xlab("Log Ita-Ub (m/s)")+ylim(c(-1,8))+theme(legend.position = "top",legend.text = element_text(size=7), axis.text = element_text(size = 7))

##Categorize distance for visualization of interaction
#distorder=c("<15km", "15-30km","30-45km",">45km")
distorder=c("<30km",">30km")##simplify visualization and easier interpretation with time of survey in other models

##Distance and nc-Ub  
Ac%<>%mutate(mindis=NA)%>%
  mutate(mindis=ifelse(exp(ndist) <=30, "<30km",">30km"))%>%
  mutate(mindis=factor(mindis, levels=distorder))


M2=Ac%>%ggplot(aes(x=nc_ubedmean, y=square(transchange),color=mindis))+geom_point(aes(x=nc_ubedmean, y=square(transchange),color=mindis), alpha=0.4)+viridis::scale_color_viridis(discrete = TRUE, option = "D")+viridis::scale_fill_viridis(discrete = TRUE, option = "D")+ylab("Observed relative change")+theme_classic()+labs(y=expression("Relative change in"~italic("Acropora")~"cover"),x="non-cyclonic Ub")+geom_smooth(method="gam", formula = y~s(x, k=3, bs="cr"))+ylim(c(-1,8))+theme(legend.position = c(0.2,0.9),legend.text = element_text(size=8), axis.text = element_text(size = 7.5))


###--------------------------
##Fit best MOdel (BRA_nACR)
###--------------------------

top.mod.fits.aic$modname[top.mod.fits.aic$response=="Other Branching"]
#[1] log.postIta.te.ndist+nc_ubedmean     nc_ubedmean.te.ndist       log.postIta.te.ndist   

Ob=groupdat%>%filter(group=="BRA_nACR")%>%droplevels()

#Best MOd
mod.ob1=gam(transchange~te(log.postIta, ndist, k = 3, bs = c("cr", "cr"))+s(nc_ubedmean, k = 3, bs = "cr")++s(reef_name,bs='re'), data=Ob,method="REML")
summary(mod.ob1)

Ob%<>%mutate(mindis=NA)%>%
  mutate(mindis=ifelse(exp(ndist) <=30, "<30km",">30km"))%>%
  mutate(mindis=factor(mindis, levels=distorder))%>%
  mutate(time=NA)%>%
  mutate(time=ifelse((loginv(log.postIta)/30) <8,"within 8 months", 
                     ifelse((loginv(log.postIta)/30) > 12 ,"After 12 months","8-12 months")))

Ob$time=ordered(Ob$time,levels=c("within 8 months","8-12 months","After 12 months"))

M3=Ob%>%ggplot(aes(x=nc_ubedmean, y=square(transchange),color=time))+geom_point(aes(x=nc_ubedmean, y=square(transchange),color=time), alpha=0.4)+viridis::scale_color_viridis(discrete = TRUE, option = "D")+viridis::scale_fill_viridis(discrete = TRUE, option = "D")+ylab("Observed relative change")+theme_classic()+labs(y="Relative change in\n Other branching cover",x="non-cyclonic Ub")+geom_smooth(method="gam", formula = y~s(x, k=3, bs="cr"))+ylim(c(-1,4))+facet_wrap(~mindis)+theme(legend.key.size=unit(0.5, "line"),legend.position = c(0.22,0.82),legend.title=element_text(size=7.5),legend.text = element_text(size=6.5), axis.text = element_text(size = 7),axis.title = element_text(size = 10))+guides(color=guide_legend(nrow=3,byrow=TRUE))


##---------------------------------
##Fit best MOdel (PORMASS)
##---------------------------------

top.mod.fits.aic$formula[top.mod.fits.aic$response=="Porites"]##Competing models 
##unique: nc_Ub+ te(postita,ndist)

Po=groupdat%>%filter(group=="POR_MASS")%>%droplevels()

# mod.por=gam(transchange~
#       s(log.postIta, k = 3, bs = "cr")+s(nc_ubedmean, k = 3, bs = "cr") +
#        + s(ndist, k = 3, bs = "cr"), data=Po,method="REML")##compared mods with ML, for AIC

mod.por=gam(transchange~
              s(nc_ubedmean, k = 3, bs = "cr") +
              + te(log.postIta,ndist, k = 3, bs = c("cr","cr"))+s(reef_name,bs='re'), data=Po,method="REML")##Same excluding reefs


Po%<>%mutate(mindis=NA)%>%
  mutate(mindis=ifelse(exp(ndist) <=30, "<30km",">30km"))%>%
  mutate(time=ifelse((loginv(log.postIta)/30) <8,"within 8 months", 
                     ifelse((loginv(log.postIta)/30) > 12 ,"After 12 months","8-12 months")))
Po$time=ordered(Po$time,levels=c("within 8 months","8-12 months","After 12 months"))


M4=Po%>%ggplot(aes(x=nc_ubedmean, y=square(transchange),color=time))+geom_point(aes(x=nc_ubedmean, y=square(transchange),color=time), alpha=0.4)+viridis::scale_color_viridis(discrete = TRUE, option = "D")+viridis::scale_fill_viridis(discrete = TRUE, option = "D")+ylab("Observed relative change")+theme_classic()+labs(y=expression('Relative change in'~italic(Porites)~"cover"),x="non-cyclonic Ub")+geom_smooth(method="gam", formula = y~s(x, k=3, bs="cr"))+ylim(c(-1,4))+facet_wrap(~mindis)+theme(legend.key.size=unit(0.5, "line"),legend.position = c(0.22,0.82),legend.title=element_text(size=7.5),legend.text = element_text(size=6.5), axis.text = element_text(size = 7),axis.title = element_text(size = 10))+guides(color=guide_legend(nrow=3,byrow=TRUE))


###--------------------------
##Fit best MOdel (MSE)
###--------------------------

top.mod.fits.aic$modname[top.mod.fits.aic$response=="MSE"]
#log.postIta+nc_ubedmean (or interaction)--  non-random effects
#nc_ubedmean+ndist-rnd model

Mse=groupdat%>%filter(group=="MSE")%>%droplevels()

mod.ms1=gam(transchange~
              te(nc_ubedmean, ndist, k = 3, bs = c("cr", "cr"))+s(reef_name,bs='re'), data=Mse,method="REML")


Mse%<>%mutate(mindis=NA)%>%
  mutate(mindis=ifelse(exp(ndist) <=30, "<30km",">30km"))%>%##more than 12mo
  mutate(mindis=factor(mindis, levels=distorder))

M5=Mse%>%ggplot(aes(x=nc_ubedmean, y=square(transchange),color=mindis))+geom_point(aes(x=nc_ubedmean, y=square(transchange),color=mindis), alpha=0.4)+viridis::scale_color_viridis(discrete = TRUE, option = "D")+viridis::scale_fill_viridis(discrete = TRUE, option = "D")+ylab("Observed relative change")+theme_classic()+labs(y="Relative change in MSE cover",x="non-cyclonic Ub")+geom_smooth(method="gam", formula = y~s(x, k=3, bs="cr"))+ylim(c(-1,4))+theme(legend.position = c(0.2,0.9),legend.text = element_text(size=8), axis.text = element_text(size = 7.5))


ggarrange(M1,M2,M3,M4,M5, ncol=2, labels = c('a)', 'b)', 'c)','d)','e)'))

```

## Plot cover response Latitudinal variability

```{r}
#relabel for plot
groupchange$group<-as.factor(groupchange$group)
levels(groupchange$group)<-c("Acropora","Other Branching","MSE","Porites massive")


g1=ggplot(groupchange, aes(x=Rchange, y=lat,color=Rchange))+geom_point(shape=21,size=1)+
  geom_jitter(alpha=0.005, width = 0.2)+
  scale_color_gradientn(colors=c("red", "#7878E6", "blue"), breaks=c(-1,-0.5,0,1,2))+ 
  geom_vline(xintercept = 0)+ 
  facet_grid(rows=vars(group))+
  theme_light()+ylab("Latitude")+xlab("Relative change in percentage cover")+
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),panel.grid.minor.y = element_blank(),
        axis.ticks.x = element_blank(),axis.ticks.y = element_blank(), legend.position = "none")


##Average per reef for better visualization
meanchange=groupchange%>%dplyr::select(-c(transectid,stid))%>%group_by(reef_name, group)%>%  summarise(change=mean(Rchange), lat=mean(lat))

g2=ggplot(meanchange, aes(x=change, y=lat, color=change))+geom_segment(aes(x=0, xend=change, y=lat,yend=lat))+geom_point(size=3)+
  geom_vline(xintercept = 0)+
  scale_color_gradientn(colors=c("red", "#7878E6", "blue"), breaks=c(-1,-0.5,0,1,2))+ 
  facet_grid(rows=vars(group))+theme_light()+ylab("")+xlab("Relative change in percentage cover")+
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),panel.grid.minor.y = element_blank(),
        axis.ticks.x = element_blank(),axis.ticks.y = element_blank(), legend.position = "none")

##-------------
##Create Map
##-------------

land <- readOGR(dsn="C:/Users/uqccastr/Dropbox/GBRMPA_Data", layer="Great_Barrier_Reef_Features")	
crs(land)

mapchange=ita_ndf
coordinates(mapchange)=~lng+lat
#crop map area using the whole latitudinal extent of the study
sw.b=extent(mapchange@bbox[1]-1,mapchange@bbox[3]+0.5,mapchange@bbox[2]-0.5, mapchange@bbox[4]+0.5)
land1 <- gSimplify(land, tol = 0.00001)
land.sub <- crop(land1, sw.b)

#Extract only coordinates of cyclone path
trackp=track[c(3,2)]#LONG and LAT

mln=(max(ita_ndf$lng))
mlt=min(ita_ndf$lat)
mlt2=max(ita_ndf$lat)

trackp%<>%filter(LON < mln)%>%filter(LAT > mlt & LAT < mlt2)

ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
)



totmap=ggplot(ita_ndf, aes(x=lng,y=lat))+
  geom_polygon(data =land.sub,aes(group = group, x = long, y = lat),colour="grey", fill=NA,alpha=0.9)+
  geom_point(data=ita_ndf, aes(colour=Rchange), size=2,shape=21)+##changed from 4 for insets
  scale_color_gradientn(colors=c("red", "#7878E6", "blue"), breaks=c(-1,-0.5,0,1,2))+ 
  theme_classic()+
  ylab("Latitude")+xlab("Longitude")+labs(color="Relative change\nin total cover")+
  coord_equal()+
  geom_path(trackp, mapping=aes(x=LON, y=LAT), color="black", alpha=0.3, size=2, lineend="round")+##Add Ita Track
  #coord_map(orientation = c(55,0,0))+
  theme_classic()+ditch_the_axes+
  theme(legend.position= c(0.4,0.5), 
        legend.background = element_rect(fill="transparent",size=8), 
        plot.title = element_text(hjust=0.5,size = 10, face = "bold"),legend.title=element_text())

#---------
##Figure 3
#---------

ggarrange(totmap,g1,g2, ncol=3,labels= c('a)','b)','c)'))

```



##Explore Reef-level cover change per group (Appendix Figure S2)

```{r}

##Greater and more frequent losses for PORMASS, MSE, Other Branching
ggplot(groupchange, aes(x=reef_name, y=Rchange, fill=group))+geom_boxplot()+
  facet_grid(rows=vars(group))+theme(axis.text.x = element_text(angle=45,hjust=1), legend.position = "none")+geom_hline(yintercept = 0, color="red", linetype="dashed")+xlab("reef")+ylab("Relative change in % cover")+scale_fill_brewer(palette="Dark2")
#ggsave(file="Outputs/pergroupchange_figS2.jpeg",width = 7, height = 8, units = c("in"),dpi = 300)

```




